An interactive environment for mapping computational structures to FPGAs

### 2016-07-07 paper/paper.tex
Last textual fixes.

  As a proof of concept to show the effectiveness of the approach, an interactive environment for simulation and hardware generation  is being developed named SDFkit.
  Using SDFkit, CSDF graphs can be visualized and simulated in a cycle accurate manner.
  Additionally, using the code can be generated for FPGAs.

  The remainder of this paper is organized as follows:
  First, the approach is covered in Section \ref{sec:approach} after which results are covered in Section \ref{sec:results}.
  Finally, conclusion and future work are expressed in Section \ref{sec:conclusions}.

@@

  The main advantage of using computational structures is that there is no need to perform dependency analysis of for-loops.
  Since all dependencies are known, computations can be distributed over space and time by mathematical transformation rules.
  Even though no standard C for-loops are used to express repetitive computations, this should not be a problem for programmers since similar patters are also used in map-reduce applications.
  Figure \ref{fig:zipwthandfoldlstruct} shows two examples of computational structures: the higher-order function \emph{zipWith} and \emph{foldl}.
  Using \emph{zipWith} an operation is applied to pairs from two lists while \emph{foldl} reduces all elements in a list to one element.

@@

  The CSDF graph representation of this specification is depicted in Figure~\ref{fig:csdf-dotproduct}.
  Each node in the graph represents a function, where function is to be interpreted in the widest sense: we regard values simply as functions that take no argument.
  The node corresponding to the addition has a self-loop, which indicates that it is involved in a \emph{stateful} computation - the fold.
  Note that neither of the two higher-order functions that appear in the specification (fold and zipWith) are represented by nodes in the CSDF graph.
  Higher-order functions do not map to individual nodes, but rather to \emph{subgraphs}.
  In general, the CSDF graphs derived from functional specifications are \emph{hierarchical}.
  That is, one may collapse a subgraph into a single node, which makes the higher-order functions explicit.
  \begin{figure}[h!]
    \centering
    \input{csdf-dp}
    \caption{Cyclo-static dataflow graph representation of dot-product.}
    \label{fig:csdf-dotproduct}
  \end{figure}

@@

  In order to determine accurate performance metrics of an application, the simulation must be accurate.
  In SDFkit, the simulation can be used to verify functional correctness as well as performance.
  Due to a specific mapping of the application graph to hardware, a cycle accurate simulation can be performed.
  Additionally, the tokens on the edges can be changed any time allowing the user to identify bottlenecks in the graph and to tweak performance.

@@

  In this paper we presented a methodology and an application for designing hardware using a structured approach.
  In contrast to translating imperative programming constructs like for-loops to hardware, we use constructs to express the structure in a computation.
  Using these computational structures, the generation of highly parallel hardware is tremendously simplified.
  Applications are modeled using CSDF graphs which are a cycle-accurate model for the generated hardware.
  This methodology is currently being implemented in SDFkit, an interactive simulation and hardware generation environment supporting visualization and simulation of CSDF graphs.
  Additionally, \clash code can be generated for HSDF graphs.

  The next step in development is to extend the hardware code generation with support for SDF and CSDF graphs.
  Additionally, the transformations to make a tradeoff between space and time are now performed outside SDFkit.
  Integration of these transformations into SDFkit is therefore an other important step forward.
  Finally, Interesting opportunities lie in extending the backend with a wider range of supported platform like C code and OpenCL.

>>>>

  As a proof of concept to show the effectiveness of the approach, an interactive environment for simulation and hardware generation  is being developed named SDFkit.
  Using SDFkit, CSDF graphs can be visualized and simulated in a cycle accurate manner.
  Additionally, code can be generated for FPGAs.

  The remainder of this paper is organized as follows:
  First, the approach is covered in Section \ref{sec:approach} after which results are covered in Section \ref{sec:results}.
  Finally, conclusions and future work are expressed in Section \ref{sec:conclusions}.

@@

  The main advantage of using computational structures is that there is no need to perform dependency analysis of for-loops.
  Since all dependencies are known, computations can be distributed over space and time by mathematical transformation rules.
  Even though no for-loops are used to express repetitive computations, this should not be a problem for programmers since similar patters are also used in map-reduce applications.
  Figure \ref{fig:zipwthandfoldlstruct} shows two examples of computational structures: the higher-order function \emph{zipWith} and \emph{foldl}.
  Using \emph{zipWith} an operation is applied to pairs from two lists while \emph{foldl} reduces all elements in a list to one element.

@@

  The CSDF graph representation of this specification is depicted in Figure~\ref{fig:csdf-dotproduct}.
  Each node in the graph represents a function, where function is to be interpreted in the widest sense: we regard values simply as functions that take no argument.
  The node corresponding to the addition has a self-loop, which indicates that it is involved in a \emph{stateful} computation - the fold.
  Note that neither of the two higher-order functions that appear in the specification (fold and zipWith) are represented by nodes in the CSDF graph.
  Higher-order functions do not map to individual nodes, but rather to \emph{subgraphs}.
  In general, the CSDF graphs derived from functional specifications are \emph{hierarchical}.
  That is, one may collapse a subgraph into a single node, which makes the higher-order functions explicit.

  \begin{figure}[h!]
    \centering
    \input{csdf-dp}
    \caption{Cyclo-static dataflow graph representation of dot-product.}
    \label{fig:csdf-dotproduct}
  \end{figure}

@@

  In order to determine performance metrics of an application, the simulation must be accurate.
  In SDFkit, the simulation can be used to verify functional correctness as well as performance.
  Due to a specific mapping of the application graph to hardware, a cycle accurate simulation can be performed.
  Additionally, the tokens on the edges can be changed any time allowing the user to identify bottlenecks in the graph and to tweak performance.

@@

  In this paper we presented a methodology and an application for designing hardware using a structured approach.
  In contrast to translating imperative programming constructs like for-loops to hardware, we use constructs to express the structure in a computation.
  Using these computational structures, the generation of highly parallel hardware is tremendously simplified.
  Applications are modeled using CSDF graphs which are a cycle-accurate model for the generated hardware.
  This methodology is currently being implemented in SDFkit, an interactive simulation and hardware generation environment supporting visualization and simulation of CSDF graphs.
  Additionally, \clash\ code can be generated for HSDF graphs.

  The next step in development is to extend the hardware code generation with support for CSDF graphs.
  Additionally, the transformations to make a tradeoff between space and time are now performed outside SDFkit.
  Integration of these transformations into SDFkit is therefore an other important step forward.
  Finally, interesting opportunities lie in extending the code generation with a wider range of supported platform like C code and OpenCL.
